### 2023-07-10 content/submission/plos_revision1/answer_to_reviewers.md

As stated above and in the text, we acquired 362 single tomographic scans of 129 different specimen.

> The first sentence of the last paragraph of the subsection Micro-computed tomographic imaging is unclear: "While performing the work, a subset of the data was always present on the production system, for working with it (see Preparation for analysis below)".
> Please provide more precise information.

We are sorry for the imprecise description of the data archival and copying process.
The last paragraph of [this section](https://habi.github.io/EAWAG-manuscript/#micro-computed-tomographic-imaging) was expanded and hopefully improved.

> The authors should revise the entire "Data analysis" subsection to present a more didactic version that helps any reader use the notebooks.
> Currently, it seems that only Python users with some expertise can understand what needs to be done.
> Perhaps the authors could provide a guide or tutorial in the GitHub repository.

The notebooks are written in Python, and are presented as Jupyter notebooks.
This makes them accessible 'piecewise', e.g. each part of the code is didactically separate and somehow compartmentalized.
The code is written in Python, so some experience with it is beneficial to thoroughly understand it.
The code and notebooks are also prepared in a way that *no* installation of *any* software except an up-to date web browser is necessary to test it.
This testing can be done by clicking one button in the code repository (linking to https://mybinder.org/v2/gh/habi/eawag/HEAD) to launch a Jupyter environment in the cloud, which is also already explained in the last paragraph before the [Discussion section](https://habi.github.io/EAWAG-manuscript/#discussion).
We understand the concerns of the reviewer though, and added explanation text and comments to the notebooks guiding (first time) users of the notebooks.
All these updates happen in [the code repository](https://github.com/habi/EAWAG) and have *not* been specifically mentioned in the manuscript text.
Two of the notebooks ([DataWrangling](https://nbviewer.org/github/habi/EAWAG/blob/main/DataWrangling.ipynb) and [DataSize](https://nbviewer.org/github/habi/EAWAG/blob/main/DataSize.ipynb)) were used to help us generate 'checking' and 'collaboration' files (DataWrangling) or used to help us generate text for the manuscript, i.e. to extract exact details on sample numbers, total amount of projections and reconstructions and their exact size on disk (DataSize).
As these notebooks are not explicitly mentioned in the manuscript they are not extensively commented (only commented).
The two notebooks mentioned in the text ([DisplayFishes](https://nbviewer.org/github/habi/EAWAG/blob/main/DisplayFishes.ipynb)) and ([ExtractOtoliths](https://nbviewer.org/github/habi/EAWAG/blob/main/ExtractOtoliths.ipynb)) have been.

@@

> How was the sanity check performed (as mentioned in the second paragraph of the subsection "Preparation for Analysis")?

The mentioned 'sanity check' was performed manually, e.g. by looking at all the values (mostly the reconstruction parameters), to exclude any operator error.
If errors were spotted at this stage, the reconstructions were deleted and re-done with the correct parameters.
Sometimes errors or problems were only spotted at a later stage, when either the pharyngeal or oral jaws were extracted, or when we noticed that some part of the the whole skull were out of the field of view of the tomographic scan prior to performing the PCA analysis mentioned later in the manuscript.
This means that specifying this as 'sanity check' is more of a glorified wording of telling the readers that we manually looked at the values of each performed tomographic scan to exlude and catch operator errors.
We reworded the sentence to hopefully make this more clear.

### 2023-07-10 manuscript.tex
Explain the notebooks a bit more. [ci skip]
This build is based on
https://github.com/habi/EAWAG-manuscript/commit/0b19da53fc1e2392fe85070eadd10f8795fff509.

This commit was created by the following CI build and job:
https://github.com/habi/EAWAG-manuscript/commit/0b19da53fc1e2392fe85070eadd10f8795fff509/checks
https://github.com/habi/EAWAG-manuscript/actions/runs/5508263410

\emph{A preprint of this manuscript is available at bioRχiv with the \href{https://doi.org/10.1101/2023.03.30.534917}{doi:10.1101/2023.03.30.534917}.}
\emph{We submitted \href{https://github.com/habi/EAWAG-manuscript/releases/tag/v1.0}{\texttt{v1.0}} (\href{https://habi.github.io/EAWAG-manuscript/v/e1e2ef76a476174a4115937d77457037ddec95df/}{online version}) of the manuscript to \href{https://journals.plos.org/plosone/}{PLOS ONE}.}
\emph{The version you see here \href{https://github.com/habi/EAWAG-manuscript/compare/v1.0...cf174315484443289e3b288555f8ec832a6cc151}{has been updated} since submission.}

This manuscript
(\href{https://habi.github.io/EAWAG-manuscript/v/cf174315484443289e3b288555f8ec832a6cc151/}{permalink})
was automatically generated
from \href{https://github.com/habi/EAWAG-manuscript/tree/cf174315484443289e3b288555f8ec832a6cc151}{habi/EAWAG-manuscript@cf17431}
on July 10, 2023.

@@

We wrote a set of \emph{Jupyter} {[}\protect\hyperlink{ref-pQ6Wbz73}{15}{]} notebooks with \emph{Python} code to work with the images and wrangle the acquired data.
The notebooks were written at the start of the project, to be able to process new scans as soon as they were reconstructed.
Re-runs of the notebook added newly scanned and reconstructed data to the analysis, facilitating an efficient quality check of the scans and batched processing of the data.

All analysis notebooks for this work are available online {[}\protect\hyperlink{ref-1HteOscVd}{16}{]}.

@@

The \href{https://github.com/habi/EAWAG/blob/master/DisplayFishes.ipynb}{main \emph{Jupyter} notebook} for this manuscript dealt with reading all log and image files and preparing images for quality checking and further analysis.

At first, \emph{all} log files of \emph{all} the data present in the processed folder were read into a \emph{pandas} {[}\protect\hyperlink{ref-19qQOHlFN}{17},\protect\hyperlink{ref-U5AcoDOX}{18}{]} dataframe.
This already enabled us to extract the specimen name and scan, since we performed multiple scans for each specimen, i.e.~a low resolution scan with large field of view for the whole head and one or two scans in high resolution focusing on the region of the oral and pharyngeal jaws.
From the log files we extracted the relevant values for double-checking the necessary parameters of each scan.
All relevant values for each scan were also saved into the dataframe and saved out to the aforementioned table in the \protect\hyperlink{supplementary-materials}{Supplementary Materials} at the end of each run of the notebook.

@@

At the end of the notebook we performed a final sanity check on the MIP images.
In this check we examined the mapping of the gray values of the raw projection images to gray values in the reconstructions, i.e.~checked that no overexposed pixels are present in the MIP images.
This is an efficient way for double-checking the gray value mapping, since the MIP images have already been generated in prior steps of the notebook and contain the highest gray values present in all the reconstructed images of each scan.

@@

Processed regions of interest were exported as \texttt{Nrrd} {[}\protect\hyperlink{ref-rn6XbmFW}{22}{]} files.
The three-dimensional visualizations of all regions of interest for each specimen were compiled into overview images (see Figure \ref{fig:104016} for an example from the compilation document).
In total we compiled overview of 125 specimens with full head morphology, oral jaw and lower pharyngeal jaw profiles.

>>>>

\emph{A preprint of this manuscript is available at bioRχiv with the \href{https://doi.org/10.1101/2023.03.30.534917}{doi:10.1101/2023.03.30.534917}.}
\emph{We submitted \href{https://github.com/habi/EAWAG-manuscript/releases/tag/v1.0}{\texttt{v1.0}} (\href{https://habi.github.io/EAWAG-manuscript/v/e1e2ef76a476174a4115937d77457037ddec95df/}{online version}) of the manuscript to \href{https://journals.plos.org/plosone/}{PLOS ONE}.}
\emph{The version you see here \href{https://github.com/habi/EAWAG-manuscript/compare/v1.0...0b19da53fc1e2392fe85070eadd10f8795fff509}{has been updated} since submission.}

This manuscript
(\href{https://habi.github.io/EAWAG-manuscript/v/0b19da53fc1e2392fe85070eadd10f8795fff509/}{permalink})
was automatically generated
from \href{https://github.com/habi/EAWAG-manuscript/tree/0b19da53fc1e2392fe85070eadd10f8795fff509}{habi/EAWAG-manuscript@0b19da5}
on July 10, 2023.

@@

We wrote a set of \emph{Jupyter} {[}\protect\hyperlink{ref-pQ6Wbz73}{15}{]} notebooks with \emph{Python} code to work with the acquired images to extract the desired data.
The notebooks were written at the start of the project, to be able to process new scans as soon as they were reconstructed.
Re-runs of the notebook added newly scanned and reconstructed data to the analysis, facilitating an efficient quality check of the scans and batched processing of the data.

All notebooks written for this work are available online {[}\protect\hyperlink{ref-1HteOscVd}{16}{]} and are extensively commented.
They are split into different task performed for this study and mentioned in their respective sections below.

@@

The \href{https://github.com/habi/EAWAG/blob/master/DisplayFishes.ipynb}{main \emph{Jupyter} notebook} for this manuscript dealt with reading all log and image files and preparing overview images of each scan for quality checking and further analysis.

At first, \emph{all} log files of \emph{all} the data present in the processed folder were read into a \emph{pandas} {[}\protect\hyperlink{ref-19qQOHlFN}{17},\protect\hyperlink{ref-U5AcoDOX}{18}{]} dataframe.
This already enabled us to extract the specimen name and scan.
Since we performed multiple scans for each specimen, i.e.~a low resolution scan with large field of view for the whole head and one or two scans in high resolution focusing on the region of the oral and pharyngeal jaws it was necessary to be able to efficiently find the scan to be looked at in detail.
From the log files we extracted the scanning and reconstruction parameters of each performed scan for double-checking them and excluding (or necessarily correcting) any operator-error.
All relevant values for each scan were also saved into the aforementioned ataframe.
This allows for an easy collation of all the relevant values into a table (as shown in the \protect\hyperlink{supplementary-materials}{Supplementary Materials}) at the end of each run of the notebook.

@@

At the end of the notebook we performed a final sanity check on the MIP images.
In this check we examined the mapping of the gray values of the raw projection images to gray values in the reconstructions, i.e.~checked that no overexposed pixels were present.
This is an efficient way for double-checking the gray value mapping, since the MIP images have already been generated in prior steps of the notebook and contain the highest gray values present in all the reconstructed images for each scan.

@@

Processed regions of interest were exported as \texttt{Nrrd} {[}\protect\hyperlink{ref-rn6XbmFW}{22}{]} files.
The three-dimensional visualizations of all regions of interest for each specimen were compiled into overview images (see Figure \ref{fig:104016} for an example from the compilation document).
In total, we compiled such overview images for 129 specimens with full head morphology, oral jaw and lower pharyngeal jaw profiles.
